---
title: "Team 1 ML project"
output: html_document
---

The goal is to group the data points into k clusters. We want variability of observations within clusters is small while variability of observations between clusters is large. 

 -- Initialization:

Randomly select k points in a feature space, these points will be our initial clusters. 

-- Iteration:

Step 1: Assign each observation to the cluster with the closest center in terms of Eucledian norm.

Step 2: Recalculate the center of each the k clusters.

Step 3. Repeat Step 1 and Step 2 until cluster assignment does not change. 

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(rpart)
library(rpart.plot)
library(caret)
library(e1071)
```

Here we are reading the Parkinson data.
```{r}
dat <- read.csv("parkinsons.csv")
```

We used MDVP.Fo.Hz., MDVP.Fhi.Hz., MDVP.Flo.Hz. variables as the first step of our cluster analysis.
 
```{r}
cluster.vars <- dat[ , c("MDVP.Fo.Hz.", "MDVP.Fhi.Hz.", "MDVP.Flo.Hz.")]
dat$MDVP.Fo.Hz. <- scale(dat$MDVP.Fo.Hz.)
dat$MDVP.Fhi.Hz. <- scale(dat$MDVP.Fhi.Hz.)
dat$MDVP.Flo.Hz. <- scale(dat$MDVP.Flo.Hz.)
```


By experimenting we found 4 to be the optimal number of centers. 
```{r}
km <- kmeans(cluster.vars, centers = 4, nstart = 50)
km
```

We created a new feature called group1 by using the clusters above.

```{r}
dat$group1 <- factor(km$cluster)
```

Here we are deleting original three features used to create a "group1". 

```{r}
dat$MDVP.Fo.Hz. <- NULL
dat$MDVP.Fhi.Hz. <- NULL
dat$MDVP.Flo.Hz. <- NULL
```

We decided that one group is not enough so we formed another cluster using MDVP.Jitter..., MDVP.Jitter.Abs., MDVP.RAP, MDVP.PPQ and Jitter.DDP variables. 

```{r}
cluster.vars2 <- dat[, c("MDVP.Jitter...", "MDVP.Jitter.Abs.", "MDVP.RAP", "MDVP.PPQ", "Jitter.DDP")]

```

By experimenting we found 4 to be the optimal number of centers. 

```{r}
km1 <- kmeans(cluster.vars2, centers = 4, nstart = 50)
km1
```

We created a new feature called group1 by using the clusters above.

```{r}
dat$group2 <- factor(km1$cluster)
```

We are deleting original features used to create the variable "group2".

```{r}
dat$MDVP.Jitter...<- NULL
dat$MDVP.Jitter.Abs. <- NULL
dat$MDVP.RAP <- NULL
dat$MDVP.PPQ <- NULL
dat$Jitter.DDP <- NULL
```

We repeated the same steps as above to create a shimmer group. 
```{r}
cluster.vars3 <- dat[, c("MDVP.Shimmer", "MDVP.Shimmer.dB.", "Shimmer.APQ3", "Shimmer.APQ5", "MDVP.APQ", "Shimmer.DDA")]
```

```{r}
km3 <- kmeans(cluster.vars3, centers = 3, nstart = 50)
km3
```

```{r}
dat$shimmer <- factor(km3$cluster)
dat$MDVP.Shimmer <- NULL
dat$MDVP.Shimmer.dB. <- NULL
dat$Shimmer.APQ3 <- NULL
dat$MDVP.APQ <- NULL
dat$Shimmer.APQ5 <- NULL
dat$Shimmer.DDA <- NULL
```

Same analysis as above to create "nmeasure" variable. 

```{r}
vars <- dat[, c("PPE", "spread1", "spread2")]
km4 <- kmeans(vars, centers = 3, nstart = 50)
km4
```


```{r}
dat$nmeasure <- factor(km4$cluster)
```

```{r}
dat$PPE <- NULL
dat$spread1 <- NULL
dat$spread2 <- NULL
```


In the following chunk we are dividing our dataset into a training and testing sets, such that the training set has 75% of the data while the testing set has the remaining 25%. 

```{r}

set.seed(1)
partition <- createDataPartition(dat$status, p = 0.75, list = FALSE)
train <- dat[partition, ]
test <- dat[-partition, ]
mean(train$status)
mean(test$status)
```

We constructed a decision tree using all features.

```{r}
tree1 <- rpart(status ~ ., data = train, method = "class", control =   rpart.control(cp = 0.0001, maxdepth = 30))
tree1
rpart.plot(tree1)
```


We tested the model on the testing set, and constructed a confusion matrix. One can see that prediction accuracy of the model is 0.875 which is very good. 

```{r}
pred <- predict(tree1, newdata = test)
pred.DT <- ifelse(pred[,2] > 0.9, "1", "0")
# saving predicted vector as factor 
Pred <- as.factor(pred.DT)
# ordering the vectors
Predicted <- ordered(Pred, levels = c("1", "0"))
Actual <- ordered(test$status,levels = c("1", "0"))
# making confusion matrix
cm <-confusionMatrix(table(Predicted,Actual))
cm
```



