---
title: "Aids Data Set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

We decided to answer Question 2: HOW MUCH DO THEY DONATE? How much do countries donate and receive? Who donates the most/least? Are there countries that donate and also receive? How does the amount donated/received by country change over time?

```{r, echo = FALSE}
aid <- read.csv(file="aiddata-countries-only.csv")
```

This next table displays the amount donated by country.

```{r, message=FALSE}
aid %>%
  group_by(donor) %>%
  summarise(total_donation=sum(commitment_amount_usd_constant)) %>% 
  slice(1:10)
```

```{r, message=FALSE}
aid %>%
  group_by(donor) %>%
  summarise(total_donation=sum(commitment_amount_usd_constant)) %>%
  arrange(desc(total_donation)) %>%
  slice(1:5) %>% 
  ggplot()+
   geom_bar(mapping=aes(x=donor, y=total_donation, fill=donor ), stat="identity") +
   labs(title="Top five donors", x="Donor", y="Total donation")
```

Here are the bottom five donors.

```{r, message=FALSE}
aid %>%
  group_by(donor) %>%
  summarise(total_donation=sum(commitment_amount_usd_constant)) %>%
  arrange(total_donation) %>%
  slice(1:5)%>% 
  ggplot()+
   geom_bar(mapping=aes(x=donor, y=total_donation, fill=donor ), stat="identity") +
   labs(title="Bottom five donors", x="Donor", y="Total donation")
```

Similarly, this table displays the amount received by country.

```{r, message = FALSE}
aid %>%
  group_by(recipient) %>%
  summarise(total_receipt=sum(commitment_amount_usd_constant)) %>% 
  slice(1:10)
```

Five top recipients.

```{r, message=FALSE}
aid %>%
  group_by(recipient) %>%
  summarise(total_receipt=sum(commitment_amount_usd_constant)) %>%
  arrange(desc(total_receipt)) %>%
  slice(1:5)%>% 
  ggplot()+
   geom_bar(mapping=aes(x=recipient, y=total_receipt, fill=recipient ), stat="identity") +
   labs(title="Top five recipients", x="Recipient", y="Total receipt")
```

Five bottom recipients.

```{r, message=FALSE}
aid %>%
  group_by(recipient) %>%
  summarise(total_receipt=sum(commitment_amount_usd_constant)) %>%
  arrange(total_receipt) %>%
  slice(1:5)%>% 
  ggplot()+
   geom_bar(mapping=aes(x=recipient, y=total_receipt, fill=recipient ), stat="identity") +
   labs(title="Bottom five recipients", x="Recipient", y="Total receipt")
```

These above tables answer Question 1 and 2. 

Question 3: which countries both receive and donate?
For Question 3, we made two dataframes, one containing distinct donors and another with distinct recipients. We used inner join on the donors and recipients tables to solve this problem.

```{r, results='hide'}
donors <- aid %>%
  distinct(donor)
donors
```

```{r, results='hide'}
recipients <- aid %>%
  distinct(recipient)
recipients
```


```{r}
donors %>% inner_join(recipients, by=c("donor"="recipient")) %>% 
  slice(1:20)
```

To address the last question, we first considered just Australia and tracked the amount donated and received over time.

```{r, results='hide', message=FALSE}
aus_donate <- aid %>%
  filter(donor =="Australia") %>%
  group_by(year) %>%
  summarise(total_donation=sum(commitment_amount_usd_constant))
aus_donate
```

```{r, results='hide', message=FALSE}
aus_receipt <- aid %>%
  filter(recipient =="Australia") %>%
  group_by(year) %>%
  summarise(total_receipt=sum(commitment_amount_usd_constant))
aus_receipt
```

```{r}
aus_donate %>% 
  left_join(aus_receipt, by = "year") %>%
  ggplot()+
  geom_line(mapping=aes(x=year, y=total_donation, color="donor"), na.rm=TRUE)+
  geom_line(mapping=aes(x=year, y=total_receipt, color="receipt"), na.rm=TRUE) +
  labs(title = "Donation and receipt from Australia", x="Year", y="Amount") 
  
```

However, we decided that it would be best to look at total donations for all countries over time.

```{r, message=FALSE}
aid %>%
  group_by(year) %>%
  summarise(total_donation_1=sum(commitment_amount_usd_constant)) %>%
  ggplot(mapping=aes(x=year, y=total_donation_1))+
  geom_line() +
  labs(title="Total donations over the year", x="Year", y="Total donation")
```

Alternatively...

```{r, message=FALSE}
aid %>%
  group_by(donor, year) %>%
  summarize(tot_don = sum(commitment_amount_usd_constant)) %>%
  filter(donor %in% c("United States", "France", "United Kingdom", "Germany", "Japan")) %>%
 ggplot() +
    geom_line(mapping = aes(year, tot_don, color = donor)) +
    labs(title="Donations of top five countries", x="Year", y="Total donation")
```

Why is there an increase around 1991? Maybe fewer countries donated?

```{r}
aid %>%
  filter(year==1991) %>%
  distinct(donor)
```
