---
title: "Predicting book success"
author: "Samal Abdikerimova, Ravi Donepudi, Carlos Salinas, and Xiran Yu"
date: today
output:
  html_document:
    toc: yes
    toc_depth: 2
    theme: united
    highlight: default
    self_contained: yes
---

# Book review statistics

```{r setup, include = FALSE}
library("broom")
library("forcats")
library("tidyverse")
library("gganimate")
```

[rgoodreads]: https://github.com/Famguy/rgoodreads
[goodreads-kaggle]: https://www.kaggle.com/jealousleopard/goodreadsbooks/
# About the data
There is a Goodreads API and an R interface to accompany it, however we will be using the data set found on Kaggle [here][goodreads-kaggle] because it is proving far too difficult to gather and process data from the Goodreads API.

First we will import the data and take a look at its contents. 
```{r import-books_csv, message = FALSE, warning = FALSE}
books <- read_csv("data/books.csv") %>%
    rename(rating = average_rating,
           total_ratings = ratings_count,
           lang = language_code,
           pages = num_pages) %>%
    na.omit()
books
```

So it looks like we are working with `r ncol(books)` variables, and `r nrow(books)` observations. Some of these variables are redundant and, for our purposes, can be ignored.

```{r }
books$bookID <- NULL
books$isbn <- NULL
books$isbn13 <- NULL
```

## Publication language
Since we speak English we'd like to focus solely on English language books. But first let's see how many of them we have

```{r, message = FALSE, warning = FALSE}
books %>%
    subset(is.na(as.numeric(lang))) %>%
    subset(!is.na(lang)) %>%
    group_by(lang) %>%
    tally() %>%
    arrange(desc(n)) %>%
    slice(1:10) %>%
    ggplot(aes(x = factor(lang, levels = lang), y = n, fill = lang)) +
    geom_bar(stat = "identity") +
    labs(title = "Title number of books by language",
         x = "language",
         y = "number of books") +
    theme(axis.text.x = element_text(angle=90, hjust=1))
```

So the overwhelming majority of the data is in `eng` so we don't loose much by ignoring the other publications. However, it looks like US and Great Britain version of books fall under a different language tag (and if we search hard Canadian too). For our purposes `"en-CA", `"en-GB"`, and `en-US` are the same so let us clean up the data a little 

```{r eng_replace, eval = FALSE}
books$lang <- books %>%
    pull(lang) %>%
    str_replace_all("en*(g|-GB|-US|-CA)", "eng")
```

One thing we might one to look at first is the distribution of ratings for English books. We do this as follows

```{r plot ratings, message = FALSE, warning = FALSE}
eng_books %>%
    group_by(rating) %>%
    summarize(total_books = n()) %>%
    ungroup() %>%
    ggplot(aes(x = rating)) +
    geom_bar(aes(y = total_books/1.000001, fill = rating),
             stat = "identity") +
    labs(title = "Total number of books per rating",
         x = "average rating",
         y = "number of books")
```

OK, to make better sense of this data let's take a look at how many people are reading these books

and who are the most prolific writers

```{r prolific writers, message = FALSE, warning = FALSE}
prolific_writer_names <- eng_books %>%
    group_by(authors) %>%
    summarize(count = n()) %>%
    arrange(desc(count)) %>%
    ungroup() %>%
    slice(1:10) %>%
    pull(authors)

prolific_writers <- eng_books %>% filter(authors %in% prolific_writer_names)
prolific_writers
```

Stephen King—that was sort of expected—and who are the top rated authors

```{r naive top authors, message = FALSE, warning = FALSE}
eng_books %>%
    group_by(authors) %>%
    summarize(average_rating = sum(rating)/n()) %>%
    arrange(desc(average_rating)) %>%
    ungroup() %>%
    slice(1:10)
```

who are these people? So highly rated authors are not necessarily notable ones, perhaps there is a better way to measure top authors.

```{r good books, message = FALSE, warning = FALSE}
eng_books %>%
    (function(...) {
        tmp <- tibble(...)
        tmp$authors <- tmp$authors %>%
            str_replace_all("Margaret Weis/Tracy Hickman",
                            "Weis/Hickman")
        tmp}) %>%
    group_by(authors) %>%
    summarize(book_count = n(),
              sum_rating = sum(rating)) %>%
    arrange(desc(sum_rating)) %>%
    slice(1:10) %>%
    ggplot(aes(factor(authors, levels = authors), sum_rating, fill = authors)) +
    geom_col() +
    coord_flip() +
    labs(title = "Author rankings",
         y = "author",
         x = "sum rating") +
    scale_fill_brewer(palette="Spectral")
```

```{r average rating, message = FALSE, rating = FALSE}
prolific_writers %>%
    (function(...) {
        tmp <- tibble(...)
        tmp$authors <- tmp$authors %>%
            str_replace_all("Margaret Weis/Tracy Hickman",
                            "Weis/Hickman")
        tmp}) %>%
    group_by(authors) %>%
    summarize(avg_rat = sum(rating)/n()) %>%
    ungroup() %>%
    ggplot(aes(x = authors, y = avg_rat, fill = authors)) +
    geom_bar(stat = "identity") +
    labs(title = "Top 10 average author rating",
          x = "auhor",
          y = "average rating") +
    theme(axis.text.x = element_text(angle = 90))
```

```{r, message = FALSE, warning = FALSE}
eng_books <- eng_books %>%
    mutate(publication_year = as.numeric(str_sub(publication_date, -4)))
```

```{r publications per year, message = FALSE, warning = FALSE}
eng_books %>% group_by(publication_year) %>%
    summarize(total_books = n()) %>%
    ungroup() %>%
    ggplot(aes(x = publication_year, y = total_books)) +
    geom_point() +
    xlim(1900, 2020) +
    labs(x = "publication year",
         y = "books published")
```

```{r, message = FALSE, rating = FALSE}
books_per_year <- eng_books %>% group_by(publication_year) %>%
    summarize(total_books = n()) %>%
    arrange(desc(total_books)) %>%
    ungroup()
```

```{r, warning = FALSE, rating = FALSE}
model_publish_books <- books_per_year %>%
    filter(publication_year <= 2006) %>%
    lm(log(total_books) ~ publication_year, .)

filtered_model <- books_per_year %>%
    filter(publication_year > 2006) %>%
    lm(log(total_books) ~ publication_year, .)

model_publish_books %>% tidy()
filtered_model %>% tidy()

estimate_2020 <-
    predict(model_publish_books,
            newdata = data.frame(publication_year = seq(2015,2020)))
```

```{r, message = FALSE, rating = FALSE}
stephen_king <- books %>% filter(authors == "Stephen King" &
                                 title != "Blood and Smoke" &
                                 title != "LT's Theory of Pets")
sk_model <- lm(rating ~ pages, stephen_king)
sk_model %>% tidy()

predict(sk_model,
        newdata = data.frame(pages = c(300)))
```

```{r, message = FALSE, rating = FALSE}
stephen_king %>% summarize(mean = mean(pages))
```

```{r, message = FALSE, warning = FALSE}
eng_books$publisher <- factor(eng_books$publisher)
```

```{r, message = FALSE, warning = FALSE }
top_10_publishers <- eng_books %>%
    group_by(publisher) %>%
    summarize(total_books = n()) %>%
    arrange(desc(total_books)) %>%
    slice(1:10) %>%
    pull(publisher)
```

```{r, message = FALSE, warning = FALSE}
eng_books_tp <- eng_books %>%
    filter(publisher %in% top_10_publishers)
```

```{r , message = FALSE, warning = FALSE}
eng_books_tp %>%
    filter(publication_year >= 2000 &
           publication_year < 2007) %>%
    ggplot(aes(x = fct_reorder(publisher, rating),
               y = rating, color = publisher)) +
    geom_boxplot() +
    labs(title = "Publisher rating",
         x = "publisher") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    transition_states(publication_year,
                      transition_length = 2,
                      state_length = 4) +
    enter_fade() +
    exit_shrink() +
    ease_aes('sine-in-out')
```

Two publishers stand out from among the rest, VIZ Media LLC and Harper Collins. VIZ media publishes graphic novels and comic books which have a higher average rating than books. 

```{r number of books}
eng_books_tp %>% filter(publisher == "Grand Central Publishing") %>%
    summarize(total = n())
```

```{r }
eng_books_tp %>% group_by(publisher) %>%
    summarize(n = n(), prop = mean(rating))
```
